### 2023-03-26 vesicle_code/docs/confinedMulti/multicomponent_vesicles.tex
working on paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Constitutive equations}
Using the same model as~\citet{liu-mar-li-vee-low2017}
and~\citet{soh-tse-li-voi-low2010}, the membrane forces depend on the
Helfrich energy, line tension, and phase energy. The phase and bending
energies are coupled through the bending modulus. In particular, the
individual energies are
\begin{align}
  E_b &= \frac{1}{2}\int_{\gamma} b(u) \kappa^2 \, ds, \\
  E_t &= \int_{\gamma} \sigma \, ds, \\
  E_p &= \frac{a}{\epsilon}\int_{\gamma}\left(
  f(u) +\frac{\epsilon^2}{2}|\nabla_\gamma u|^2\right) \, ds,
  \label{eqn:PhaseEnergy}
\end{align}
where $u$ is the dimensionless lipid concentration, $b(u)$ is the
lipid-dependent bending modulus, $\sigma$ is the vesicle's tension, and
$f(u) = \frac{1}{4}u^2(1-u)^2$ is the double-well potential with local
minimums at $u=0$ and $u=1$. Here, $\kappa$ is the membrane curvature,
$\ss$ is the tangent vector, and $s$ is the arclength. We have adopted
the same setup as~\citet{soh-tse-li-voi-low2010} where $\epsilon \ll 1$
is a dimensionless parameter that controls the size of the transition
region of $u$, and $a$ is the dimensional line tension scaled by
$b_{\max}/L$. Both $\epsilon$ and $a$ are kept constant in all
simulations. The resulting membrane forces are
\begin{align}
  \ff_b &= -(b(u)\kappa \nn)_{ss} -\frac{3}{2}
    \left(b(u) \kappa^2 \ss\right)_s, \\
  \ff_t &= (\sigma \ss)_s, \\
  \ff_p &= \left(\frac{a}{\epsilon}\left(f(u) -
     \frac{\epsilon^2}{2} u_s^2\right) \ss \right)_s.
\end{align}

The lipid species $u$ is governed by a fourth-order Cahn-Hilliard equation that results in the lipids phase separating while conserving their total mass. To model the variable bending, \citet{soh-tse-li-voi-low2010} propose parameterizing the bending modulus as
\begin{align}
    b(u) = (1-u) + \beta u.
    \label{eqn:linearBending}
\end{align}
However, since the double-well potential does not have hard walls, the
lipid concentration is not guaranteed to be confined to the interval
$[0,1]$. This is problematic when $u > 1$ since this results in $b(u) <
0$ if $u > (1 - \beta)^{-1}$, and such values of $u$ are possible when
$\beta$ is small and the vesicle is confined to a narrow region.
Therefore, in this work we parameterize the bending modulus as
\begin{align}
    b(u) = \frac{\beta-1}{2} \tanh\left(3\left(u-\frac{1}{2} \right)\right) + 
           \frac{\beta + 1}{2}.
    \label{eqn:tanhBending}
\end{align}
\begin{figure}[H]
  \centering
  \input{figures/concModels.tikz}
  \caption{\label{fig:concModels} \small The linear bending model in red
  (equation~\eqref{eqn:linearBending}) and the the new hyperbolic
  bending model in blue (equation~\eqref{eqn:tanhBending}). Note that
  the new bending model satisfies $b \in (\beta,1)$ even if $u \notin
  [0,1]$.}
\end{figure}

Equation~\eqref{eqn:tanhBending} maps the local minimums of the
double-well potential close to $\beta$ and $1$, but more
importantly, b(u) remains bounded in $(\beta,1)$ even when $u
\notin [0,1]$ (Figure~\ref{fig:concModels}). To demonstrate the
necessity of this new bending model, we consider a multicomponent
vesicle in strong confinement with both bending
models~\eqref{eqn:linearBending} and~\eqref{eqn:tanhBending}. The
stiff-to-floppy ratio is $\beta = 10^{-1}$. In the new
model~\eqref{eqn:tanhBending}, the lipid concentration achieves a
maximum value of $u \approx 1.002$ which results in a  minimum bending
stiffness of $b(u) \approx \beta$. However, if we use the linear bending
model~\eqref{eqn:linearBending}, the lipid concentration achieves a
maximum value of $u \approx 1.52$ which results in an unphsical negative
bending stiffness of $b(u) \approx -0.5$. The vesicle shapes are in
Figure~\ref{fig:bendingModels}, and the instability caused by the
negative bending stiffness is near the region with negative curvature.
\begin{figure}[htp]
  \centering
%    \begin{tabular}{@{}p{0.49\linewidth}@{\qquad}p{0.49\linewidth}@{}}
  \subfigimg[width=0.48\linewidth, clip ]{(a)}{figures/newbending.pdf}
  \subfigimg[width=0.48\linewidth, clip ]{(b)}{figures/oldbending.pdf}
%    \end{tabular}
\caption{\label{fig:bendingModels} \small A multicomponent vesicle
  entering a closely-fitting channel using (a) the new bending
  model~\eqref{eqn:tanhBending}; (b) the linear
  model~\eqref{eqn:linearBending}. With the linear model, the vesicle
  bending modulus becomes negative, and this introduces an instability
  that first appears near the point with the largest negative curvature.
  The two plots are not at the same time step.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\label{sec:NumericalMethods}Numerical Methods}
The velocity field $\uu$ is written as a combination of a single-layer potential and a double-layer potential
\begin{align}
    \label{eqn:LPrep}
    \uu(\xx) = S[\ff](\xx) + D[\eeta](\xx),
\end{align}
where
\begin{align}
    S[\ff](\xx) &= \frac{1}{4\pi\mu} \int_{\gamma} \left(-\log \rho \mathds{I} + \frac{\rr \otimes \rr}{\rho^2} \right)
    \ff(\yy) \, ds_\yy, \\
%
    D[\eeta](\xx) &= \frac{1}{\pi} \int_{\Gamma} \frac{\rr \cdot \nn}{\rho^2}
        \frac{\rr \otimes \rr}{\rho^2} \eeta(\yy) \, ds_\yy,
\end{align}
$\rr = \xx - \yy$, $\rho = |\rr|$, and $\mathds{I}$ is the $2 \times 2$ identity matrix. To avoid tangling of the vesicle membrane, it is parameterized in terms of the $\theta$-$L$ variables~\cite{hou-low-she1994}, where $L$ is the fixed vesicle length and $\theta$ is the opening angle of the tangent vector $\ss$. This requires decomposing the velocity field on $\gamma$ into its normal direction $V$ and tangential direction $T$. Then, the no-slip boundary condition for the vesicle velocity is
\begin{align}
    \frac{d\xx}{dt} = V \nn + T \ss + D[\eeta](\xx), \quad \xx \in \gamma.
\end{align}
The boundary condition on $\Gamma$ is imposed by requiring that $\eeta$ satisfies
\begin{align}
    \UU(\xx) = -\frac{1}{2}\eeta(\xx) + S[\ff](\xx) + D[\eeta](\xx), \quad \xx \in \Gamma.
    \label{eqn:DLP_BIE}
\end{align}
After discretizing the geometry at collocation points, the single-layer potential is approximated with quadrature for weakly-singular integrands, and the double-layer potential is approximated with the spectrally accurate trapezoid rule. All arclength derivatives are computed with spectral accuracy using Fourier methods. Time stepping is performed with second-order Adams-Bashforth. The resulting linear system is solved with GMRES.

@@

To define the size of the top lubrication layer (the size of the bottom lubrication layer is defined similarly), we start with the distance between the vesicle and the top half of the confining geometry
\begin{align}
    d(\xx,\Gamma_\mathrm{top}) \inf_{\yy \in \Gamma_\mathrm{top}} \|\xx-\yy\|,
\end{align}
where $\Gamma_{\mathrm{top}}$ is the top portion of $\Gamma$. We then find all local minimums of $d(\xx,\Gamma_\mathrm{top})$, with the condition that $d(\xx,\Gamma_\mathrm{top}) < \beta$, where $\beta$ is a parameter. If no such local minimums exist, then the vesicle is too far from $\Gamma_\mathrm{top}$ to define a lubrication layer. If there are two or more local minimums, the left-most, $\xx_L$, and right-most, $\xx_R$, local minimums form the start and end of the lubrication layer. If there is only one local minimum, then $\xx_L$ and $\xx_R$ are the points on $\gamma$ to the left and right of this local minimum with the condition that $d(\xx_L,\Gamma_\mathrm{top}) = d(\xx_R,\Gamma_\mathrm{top}) = \beta$. Then, the segment $\gamma_{\mathrm{layer}} \subset \gamma$ consists of all points $\xx \in \gamma$ between $\xx_L$ and $\xx_R$ with the condition that $d(\xx,\Gamma_\mathrm{top}) < \beta$. Finally, the lubrication layer width is
\begin{align}
    w_\mathrm{top} = \frac{1}{|\gamma_{_\mathrm{layer}}|} \int_{\gamma_{_\mathrm{layer}}} d(\xx,\Gamma_\mathrm{top}) \, ds.
\end{align}
The width of the bottom lubrication layer, $w_\mathrm{bot}$ is calculated replacing $\Gamma_\mathrm{top}$ with $\Gamma_\mathrm{bot}$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\label{sec:results}Hydrodynamics of a multicomponent vesicle under strong confinement}
We consider both a single-component and a multicomponent vesicle suspended in two geometries: a closely-fitting geometry (stenosis) and a geometry that slowly contracts to a narrow neck and then quickly widens (contracting). We considered a semipermeable vesicle suspended in these same geometries~\cite{qua-gan-you2021}. All the multicomponent examples have the bending ratio $b_{\min}/b_{\max} = 10^{-1}$. For the single-component case, we consider two dimensionless bending stiffness: $b(u) = b_{\max} = 1$ and $b(u) = \frac{1}{2}(b_{\max} + b_{\min}) = 0.55$. Throughout this section we plot the position of the vesicle in terms of the $x$-coordinate of its center of mass. For the stenosed geometry in Section~\ref{subsec:Stenosis}, the narrow region begins at $x=-15$ and ends at $x=15$. For the contracting geometry in Section~\ref{subsec:Contraction}, the geometry begins to narrow at $x=4$ and reaches its narrowest point at $x=15$.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Constitutive equations}
Using the same model as~\citet{liu-mar-li-vee-low2017}
and~\citet{soh-tse-li-voi-low2010}, the membrane forces depend on the
Helfrich energy, line tension, and phase energy. The phase and bending
energies are coupled through the bending modulus. In particular, the
individual energies are
\begin{align}
  E_b &= \frac{1}{2}\int_{\gamma} b(u) \kappa^2 \, ds, \quad
  E_t = \int_{\gamma} \sigma \, ds, \\
  E_p &= \frac{a}{\epsilon}\int_{\gamma}\left(
  f(u) +\frac{\epsilon^2}{2}|\nabla_\gamma u|^2\right) \, ds,
  \label{eqn:PhaseEnergy}
\end{align}
where $u$ is the dimensionless lipid concentration, $b(u)$ is the
lipid-dependent bending modulus, $\sigma$ is the vesicle's tension, and
$f(u) = \frac{1}{4}u^2(1-u)^2$ is the double-well potential with local
minimums at $u=0$ and $u=1$. Here, $\kappa$ is the membrane curvature,
$\ss$ is the tangent vector, and $s$ is the arclength. We have adopted
the same setup as~\citet{soh-tse-li-voi-low2010} where $\epsilon \ll 1$
is a dimensionless parameter that controls the size of the transition
region of $u$, and $a$ is the dimensional line tension scaled by
$b_{\max}/L$. Both $\epsilon$ and $a$ are kept constant in all
simulations. The resulting membrane forces are
\begin{align}
  \ff_b &= -(b(u)\kappa \nn)_{ss} -\frac{3}{2}
    \left(b(u) \kappa^2 \ss\right)_s, \\
  \ff_t &= (\sigma \ss)_s, \\
  \ff_p &= \left(\frac{a}{\epsilon}\left(f(u) -
     \frac{\epsilon^2}{2} u_s^2\right) \ss \right)_s.
\end{align}

\begin{figure}[H]
  \centering
%  \input{figures/concModels.tikz} \\
  \subfigimg[width=0.9\linewidth, clip ]{(a)}{figures/concModels.pdf}
  \\
  \subfigimg[width=0.9\linewidth, clip ]{(b)}{figures/oldbending.pdf}
  \caption{\label{fig:concModels} \small (a) The linear bending model in
  red (equation~\eqref{eqn:linearBending}) and the new hyperbolic
  bending model in blue (equation~\eqref{eqn:tanhBending}). Note that
  the new bending model satisfies $b \in (\beta,1)$ even if $u \notin
  [0,1]$. (b) A multicomponent vesicle entering a closely-fitting
  channel using the linear model~\eqref{eqn:linearBending}. The vesicle
  bending modulus becomes negative, and this introduces an instability
  that first appears near the point with the largest negative curvature.
  }
\end{figure}

The lipid species $u$ is governed by a fourth-order Cahn-Hilliard equation that results in the lipids phase separating while conserving their total mass. To model the variable bending, \citet{soh-tse-li-voi-low2010} propose parameterizing the bending modulus as
\begin{align}
    b(u) = (1-u) + \beta u.
    \label{eqn:linearBending}
\end{align}
However, since the double-well potential does not have hard walls, the
lipid concentration is not guaranteed to be confined to the interval
$[0,1]$. This is problematic when $u > 1$ since this results in $b(u) <
0$ if $u > (1 - \beta)^{-1}$, and such values of $u$ are possible when
$\beta$ is small and the vesicle is confined to a narrow region.
Therefore, in this work we parameterize the bending modulus as
\begin{align}
    b(u) = \frac{\beta-1}{2} \tanh\left(3\left(u-\frac{1}{2} \right)\right) + 
           \frac{\beta + 1}{2}.
    \label{eqn:tanhBending}
\end{align}
This maps the local minimums of the
double-well potential close to $\beta$ and $1$, but more
importantly, b(u) remains bounded in $(\beta,1)$ even when $u
\notin [0,1]$ (Figure~\ref{fig:concModels}(a)). To demonstrate the
necessity of this new bending model, we consider a multicomponent
vesicle in strong confinement with both bending
models~\eqref{eqn:linearBending} and~\eqref{eqn:tanhBending}. The
stiff-to-floppy ratio is $\beta = 10^{-1}$. In the new
model~\eqref{eqn:tanhBending}, the lipid concentration achieves a
maximum value of $u \approx 1.002$ which results in a  minimum bending
stiffness of $b(u) \approx \beta$. However, if we use the linear bending
model~\eqref{eqn:linearBending}, the lipid concentration achieves a
maximum value of $u \approx 1.52$ which results in an unphysical
negative bending stiffness of $b(u) \approx -0.5$. The corresponding
vesicle shape is in Figure~\ref{fig:concModels}(b), and the instability
caused by the negative bending stiffness is near the region with
negative curvature.
%\begin{figure}[htp]
%  \centering
%%    \begin{tabular}{@{}p{0.49\linewidth}@{\qquad}p{0.49\linewidth}@{}}
%  \subfigimg[width=0.48\linewidth, clip ]{(a)}{figures/newbending.pdf}
%%    \end{tabular}
%\caption{\label{fig:bendingModels} \small A multicomponent vesicle
%  entering a closely-fitting channel using (a) the new bending
%  model~\eqref{eqn:tanhBending}; (b) the linear
%  model~\eqref{eqn:linearBending}. With the linear model, the vesicle
%  bending modulus becomes negative, and this introduces an instability
%  that first appears near the point with the largest negative curvature.
%  The two plots are not at the same time step.}
%\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\label{sec:NumericalMethods}Numerical Methods}
The velocity field $\uu$ is written as a combination of a single-layer potential and a double-layer potential
\begin{align}
    \label{eqn:LPrep}
    \uu(\xx) = S[\ff](\xx) + D[\eeta](\xx),
\end{align}
where
\begin{align}
    S[\ff](\xx) &= \frac{1}{4\pi\mu} \int_{\gamma} \left(-\log \rho \mathds{I} + \frac{\rr \otimes \rr}{\rho^2} \right)
    \ff(\yy) \, ds_\yy, \\
%
    D[\eeta](\xx) &= \frac{1}{\pi} \int_{\Gamma} \frac{\rr \cdot \nn}{\rho^2}
        \frac{\rr \otimes \rr}{\rho^2} \eeta(\yy) \, ds_\yy,
\end{align}
$\rr = \xx - \yy$, $\rho = |\rr|$, and $\mathds{I}$ is the $2 \times 2$ identity matrix. To avoid tangling of the vesicle membrane, it is parameterized in terms of the $\theta$-$L$ variables~\cite{hou-low-she1994}, where $L$ is the fixed vesicle length and $\theta$ is the opening angle of the tangent vector $\ss$. This requires decomposing the velocity field on $\gamma$ into its normal direction $V$ and tangential direction $T$. Then, the no-slip boundary condition for the vesicle velocity is
\begin{align}
    \frac{d\xx}{dt} = V \nn + T \ss + D[\eeta](\xx), \quad \xx \in \gamma.
\end{align}
The boundary condition on $\Gamma$ is imposed by requiring that $\eeta$ satisfies
\begin{align}
    \UU(\xx) = -\frac{1}{2}\eeta(\xx) + S[\ff](\xx) + D[\eeta](\xx), \quad \xx \in \Gamma.
    \label{eqn:DLP_BIE}
\end{align}
After discretizing the geometry at collocation points, the single-layer potential is approximated with quadrature for weakly-singular integrands, and the double-layer potential is approximated with the spectrally accurate trapezoid rule. All arclength derivatives are computed with spectral accuracy using Fourier methods. Time stepping is performed with second-order Adams-Bashforth. The resulting linear system is solved with GMRES.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\label{sec:results}Hydrodynamics of a multicomponent vesicle under strong confinement}

@@

To define the size of the top lubrication layer (the size of the bottom
lubrication layer is defined similarly), we let
$d(\xx,\Gamma_\mathrm{top})$ be the minimum distance between the vesicle
and the top half of the confining geometry.
%\begin{align}
%    d(\xx,\Gamma_\mathrm{top}) = \inf_{\yy \in \Gamma_\mathrm{top}} \|\xx-\yy\|,
%\end{align}
%where $\Gamma_{\mathrm{top}}$ is the top portion of $\Gamma$.
We then find all local minimums of $d(\xx,\Gamma_\mathrm{top})$, with
the condition that $d(\xx,\Gamma_\mathrm{top}) < \beta$, where $\beta$
is a parameter. If no such local minimums exist, then the vesicle is too
far from $\Gamma_\mathrm{top}$ to define a lubrication layer. If there
are two or more local minimums, the left-most, $\xx_L$, and right-most,
$\xx_R$, local minimums form the start and end of the lubrication layer.
If there is only one local minimum, then $\xx_L$ and $\xx_R$ are the
points on $\gamma$ to the left and right of this local minimum with the
condition that $d(\xx_L,\Gamma_\mathrm{top}) =
d(\xx_R,\Gamma_\mathrm{top}) = \beta$. Then, the segment
$\gamma_{\mathrm{layer}} \subset \gamma$ consists of all points $\xx \in
\gamma$ between $\xx_L$ and $\xx_R$ with the condition that
$d(\xx,\Gamma_\mathrm{top}) < \beta$. Finally, the lubrication layer
width is
\begin{align}
    w_\mathrm{top} = \frac{1}{|\gamma_{_\mathrm{layer}}|} \int_{\gamma_{_\mathrm{layer}}} d(\xx,\Gamma_\mathrm{top}) \, ds.
\end{align}
The width of the bottom lubrication layer, $w_\mathrm{bot}$ is calculated replacing $\Gamma_\mathrm{top}$ with $\Gamma_\mathrm{bot}$.

@@

We consider both a single-component and a multicomponent vesicle suspended in two geometries: a closely-fitting geometry (stenosis) and a geometry that slowly contracts to a narrow neck and then quickly widens (contracting). We considered a semipermeable vesicle suspended in these same geometries~\cite{qua-gan-you2021}. All the multicomponent examples have the bending ratio $b_{\min}/b_{\max} = 10^{-1}$. For the single-component case, we consider two dimensionless bending stiffness: $b(u) = b_{\max} = 1$ and $b(u) = \frac{1}{2}(b_{\max} + b_{\min}) = 0.55$. Throughout this section we plot the position of the vesicle in terms of the $x$-coordinate of its center of mass. For the stenosed geometry in Section~\ref{subsec:Stenosis}, the narrow region begins at $x=-15$ and ends at $x=15$. For the contracting geometry in Section~\ref{subsec:Contraction}, the geometry begins to narrow at $x=4$ and reaches its narrowest point at $x=15$.
